How crisis affect the low-income and high-income economies differently

### 2024-04-09 paper/paper.qmd
Working on Data section

library(tidyverse)
library(readr)
library(dplyr)
library(arrow)
library(dataverse)
library(WDI)

@@

data_na |>
  ggplot(mapping = aes(x = low_income, y = inflation)) +
  geom_point() +
  theme_minimal() +
  labs(x = "Low-income countries", y = "Inflation")+
  scale_x_continuous(breaks = c(0,1))
```

```{r}
#| label: fig-gdp growth
#| fig-cap: Relationship between differenct level of income countries versus gdp growth
#| echo: false
#| warning: false
#| message: false

data_na |>
  ggplot(mapping = aes(x = low_income, y = gdp_growth)) +
  geom_point() +
  theme_minimal() +
  labs(x = "Low-income countries", y = "gdp_growth")+
  scale_x_continuous(breaks = c(0,1))
```

```{r}
data_na |>
  ggplot(mapping = aes(x = low_income, y = lf_par_rate)) +
  geom_point() +
  theme_minimal() +
  labs(x = "Low-income countries", y = "Inflation")+
  scale_x_continuous(breaks = c(0,1))
```

@@

@@

# Appendix {-}

# Additional data details

>>>>

library(tidyverse)
library(readr)
library(dplyr)
library(arrow)
library(dataverse)
library(WDI)
library(rstanarm)

@@

data_na |>
  ggplot(mapping = aes(x = low_income, y = inflation)) +
  geom_point() +
  theme_minimal() +
  labs(x = "Low-income countries", y = "Inflation")+
  scale_x_continuous(breaks = c(0,1))
```
In @fig-inflation, I have shown the inflation rate for the low-income and the high-income countries. On the x-axis, when x is equal to 0, it means that the county is a high-income country and it equals to 1 when the country is a low-income country. From @fig-inflation, we could see that high income countries has lower and more stabilized inflation rates. In low-income countries the inflation rate varies across each country.

```{r}
#| label: fig-gdp-growth
#| fig-cap: Relationship between differenct level of income countries versus gdp growth
#| echo: false
#| warning: false
#| message: false

data_na |>
  ggplot(mapping = aes(x = low_income, y = gdp_growth)) +
  geom_point() +
  theme_minimal() +
  labs(x = "Low-income countries", y = "gdp_growth")+
  scale_x_continuous(breaks = c(0,1))
```
In @fig-gdp-growth, I have shown the gdp growth rate for the low-income and the high-income countries. On the x-axis, when x is equal to 0, it means that the county is a high-income country and it equals to 1 when the country is a low-income country. From @fig-gdp-growth, we could see that high income countries have similar gdp growth rates. In low-income countries the gdp growth rates varies across each country.

```{r}
#| label: fig-lf
#| fig-cap: Relationship between differenct level of income countries versus labour force participation rate
#| echo: false
#| warning: false
#| message: false
data_na |>
  ggplot(mapping = aes(x = low_income, y = lf_par_rate)) +
  geom_point() +
  theme_minimal() +
  labs(x = "Low-income countries", y = "labour force rate")+
  scale_x_continuous(breaks = c(0,1))
```

From the data presented in @fig-inflation, @fig-gdp-growth, and @fig-lf, I observed that high-income countries shared similar economic performance trends. Despite not necessarily excelling in all economic indicators than the low-income countries, however, these countries generally maintain a more stable economic condition, with only a few exceptions or outliers noted.

 A more detailed relationship between each variable could be found in @fig-relationship.

@@

```{r}

lowincome <- data_na %>% filter(low_income == 1)
highincome <- data_na %>% filter(low_income == 0)
lowmodel <- glm(gni_per_cap ~ gdp_growth + population + unem_rate + lf_par_rate + trade_per + inflation, 
             family = gaussian(link = "identity"), data = lowincome)

summary(lowmodel)

```{r}
highmodel <- glm(gni_per_cap ~ gdp_growth + population + unem_rate + lf_par_rate + trade_per + inflation, 
             family = gaussian(link = "identity"), data = highincome)

summary(highmodel)
```

@@

# Additional data details
```{r}
#| label: fig-relationship 
#| fig-cap: Scatterplot matrix of each variables
#| echo: false
#| warning: false
#| message: false

pairs(~inflation + gdp_growth + population + unem_rate + lf_par_rate + trade_per + gni_per_cap, data = data_na)
```
